---
title: "week 8 lab hw"
author: "Jeremy Wadowski"
date: "March 1, 2017"
output: github_document
---

```{r}
#http://tutorials.iq.harvard.edu/R/Rgraphics/Rgraphics.html#org93999d8
housing <- read.csv("/home/eeb177-student/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/landdata-states.csv")
head(housing[1:5])

#Base graphics histogram example
hist(housing$Home.Value)
```
## ggplot2 histogram example  
```{r}
library(ggplot2)
ggplot(housing, aes(x = Home.Value)) + geom_histogram()

```
## Base colored scatter plot example  
```{r}
plot(Home.Value ~ Date,
     data=subset(housing, State == "MA"))
points(Home.Value ~ Date, col="red",
       data=subset(housing, State == "TX"))
legend(1975, 400000,
       c("MA", "TX"), title="State",
       col=c("black", "red"),
       pch=c(1, 1))
```
## ggplot2 colored scatter plot example  
```{r}
ggplot(subset(housing, State %in% c("MA", "TX")),
       aes(x=Date,
           y=Home.Value,
           color=State))+
  geom_point()
```
# Geometric Objects and Aesthetic  
```{r}
hp2001Q1 <- subset(housing, Date == 2001.25) 
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = Land.Value)) +
  geom_point()

hp2001Q1$pred.SC <- predict(lm(Structure.Cost ~ log(Land.Value), data = hp2001Q1))

p1 <- ggplot(hp2001Q1, aes(x = log(Land.Value), y = Structure.Cost))

p1 + geom_point(aes(color = Home.Value)) +
  geom_line(aes(y = pred.SC))

p1 +
  geom_point(aes(color = Home.Value)) +
  geom_smooth()

p1 + 
  geom_text(aes(label=State), size = 3)

#install.packages("ggrepel") 
library("ggrepel")
p1 + 
  geom_point() + 
  geom_text_repel(aes(label=State), size = 3)
p1 +
  geom_point(aes(size = 2),# incorrect! 2 is not a variable
             color="red") # this is fine -- all points red
p1 +
  geom_point(aes(color=Home.Value, shape = region))

```
# Exercise 1  
```{r}
dat <- read.csv("/home/eeb177-student/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/EconomistData.csv")
head(dat)

ggplot(dat, aes(x = CPI, y = HDI, size = HDI.Rank)) + geom_point()

ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()

ggplot(dat, aes(x = CPI, y = HDI), color="blue") + geom_point()

ggplot(dat, aes(x = CPI, y = HDI, size = HDI.Rank, color=Region)) + geom_point()



```
# Setting Statistical Transformation Arguments  
```{r}
args(geom_histogram)
args(stat_bin)

p2 <- ggplot(housing, aes(x = Home.Value))
p2 + geom_histogram()

#with changed bin value
p2 + geom_histogram(stat = "bin", binwidth=4000)

#Changing the Statistical Transformation
housing.sum <- aggregate(housing["Home.Value"], housing["State"], FUN=mean)
rbind(head(housing.sum), tail(housing.sum))

#Doesn't work because we are asking ggplot to bin and summarize already binned and summarized data
#ggplot(housing.sum, aes(x=State, y=Home.Value)) + geom_bar()
#instead:
ggplot(housing.sum, aes(x=State, y=Home.Value)) + 
  geom_bar(stat="identity")
```
# Exercise 2  
```{r}
dat <- read.csv("/home/eeb177-student/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/EconomistData.csv")
head(dat)

ggplot(dat, aes(x = CPI, y = HDI))+ geom_point()

ggplot(dat, aes(x = CPI, y = HDI))+ geom_point() + geom_smooth()

c(method="lm")

ggplot(dat, aes(x = CPI, y = HDI))+ geom_point() + geom_line(stat="identity")

ggplot(dat, aes(x = CPI, y = HDI))+ geom_point() + geom_smooth(span = .2)

```
## Scale Modification Examples  
```{r}
p3 <- ggplot(housing,
             aes(x = State,
                 y = Home.Price.Index)) + 
        theme(legend.position="top",
              axis.text=element_text(size = 6))
(p4 <- p3 + geom_point(aes(color = Date),
                       alpha = 0.5,
                       size = 1.5,
                       position = position_jitter(width = 0.25, height = 0)))

p4 + scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"))



p4 +
  scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"),
                         low = "blue", high = "red")


#p4 +
  #scale_color_continuous(name="",
                         #breaks = c(1976, 1994, 2013),
                         #labels = c("'76", "'94", "'13"),
                         #low = muted("blue"), high = muted("red"))  #does not work with current package, requires scale package

#midpoint and different color scales
p4 +
  scale_color_gradient2(name="",
                        breaks = c(1976, 1994, 2013),
                        labels = c("'76", "'94", "'13"),
                        low = ("blue"),
                        high = ("red"),
                        mid = "gray60",
                        midpoint = 1994)


```
## Exercise 3  
```{r}
dat <- read.csv("/home/eeb177-student/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/EconomistData.csv")

ggplot(dat, aes(x = CPI, y = HDI, color = "Region")) +
    geom_point()

  ggplot(dat, aes(x = CPI, y = HDI, color = "Region")) +
  geom_point() +
  scale_x_continuous(name = "Corruption Perception Index") +
  scale_y_continuous(name = "Human Development Index") +
  scale_color_discrete(name = "World Regions")

 ggplot(dat, aes(x = CPI, y = HDI, color = "Region")) +
  geom_point() +
  scale_x_continuous(name = "Corruption Perception Index") +
  scale_y_continuous(name = "Human Development Index") +
    scale_color_manual(name = "Region of the world",
                       values = c("#24576D",
                                  "#099DD7",
                                  "#28AADC",
                                  "#248E84",
                                  "#F2583F",
                                  "#96503F")) #no different colors?? 
  
```
## Faceting  
```{r}
p5 <- ggplot(housing, aes(x = Date, y = Home.Value))
p5 + geom_line(aes(color = State)) #produces obscure lines

(p5 <- p5 + geom_line() +
   facet_wrap(~State, ncol = 10))

```
## Themes  
```{r}
p5 + theme_linedraw()

p5 + theme_light()

p5 + theme_minimal() +
  theme(text = element_text(color = "turquoise"))

#creating new themes
theme_new <- theme_bw() +
  theme(plot.background = element_rect(size = 1, color = "blue", fill = "black"),
        text=element_text(size = 12, family = "Serif", color = "ivory"),
        axis.text.y = element_text(colour = "purple"),
        axis.text.x = element_text(colour = "red"),
        panel.background = element_rect(fill = "pink"),
        strip.background = element_rect(fill = ("orange")))

p5 + theme_new
```
## Best way to plot a data.frame of two variables as separate points with different colors?  
```{r}

#not this way:
library(tidyr)
housing.byyear <- aggregate(cbind(Home.Value, Land.Value) ~ Date, data = housing, mean)
ggplot(housing.byyear,
       aes(x=Date)) +
  geom_line(aes(y=Home.Value), color="red") +
  geom_line(aes(y=Land.Value), color="blue")

#but this way:
home.land.byyear <- gather(housing.byyear,
                           value = "value",
                           key = "type",
                           Home.Value, Land.Value)
ggplot(home.land.byyear,
       aes(x=Date,
           y=value,
           color=type)) +
  geom_line()

```
# Putting It All Together: Recreate Economist Graph  
```{r}
dat <- read.csv("/home/eeb177-student/Desktop/eeb-177/lab-work/exercise-8/Rgraphics/dataSets/EconomistData.csv")

first <- ggplot(dat, aes(x = CPI, y = HDI, color = Region))
first + geom_point()

(second <- first + geom_smooth(aes(group = 1), method = "lm", formula = y ~ log(x), se = FALSE, color = "red")) + geom_point()

(third <- second + geom_point(size = 1, shape = 1) + geom_point(size = 1, shape = 1) + geom_point(size = 1, shape = 1))


labeled_countries <- c("Congo", "Afghanistan", "Sudan", "Myanmar", "Iraq", "India", "Rwanda", "South Africa", "China", "Venezuela", "Russia", "Argentina", "Greece", "Brazil", "Italy", "Bhutan", "Cape Verde", "Botswana", "Spain", "France", "US", "Britain", "Barbados", "Japan", "Germany", "Norway", "Singapore", "New Zealand")

dat$Country <- as.character(dat$Country)
as.character(dat$Country) %in% labeled_countries
country_to_print <- character(nrow(dat))
printrows <- which(as.character(dat$Country) %in% labeled_countries) 
#country_to_print[printrows] <- #dat$Country[printrows]

library("ggrepel")
(fourth <- third + geom_text_repel(aes(label = Country), color = "black", data = subset(dat, Country%in%labeled_countries), force = 6))

library(grid)
(fifth <- fourth + ggtitle("Corruption and human development") + scale_x_continuous(name = "Corruption Perceptions Index, 2011 (10=least corrupt)", limits = c(.9, 10.5), breaks = 1:10) + scale_y_continuous(name = "Human Development Index, 2011 (1=Best)", limits = c(0.2, 1.0), breaks = seq(0.2, 1.0, by = 0.1)))

(sixth <- fifth + theme_minimal() + theme(text = element_text(color = "black"), legend.position = c("top"), legend.direction = "horizontal", legend.justification = 0.25, legend.text = element_text(size = 8, color = "black"), axis.text = element_text(face = "italic", size = 5), axis.ticks.y = element_blank(), axis.line = element_line(color = "black", size = 0.4),axis.line.y =element_blank(), panel.grid.major = element_line(color = "black", size = 0.4),panel.grid.major.x = element_blank()))

```
